---
title: "Reproducible Research, Assignment #1"
author: "gnpitty"
date: "Saturday, December 13, 2014"
output: html_document
 
---

### Loading and preprocessing the data

```{r, echo=TRUE}
data  <- read.table("activity.csv", header = TRUE, sep = "," ) 
library(data.table)
library (ggplot2)
library (plyr)
```  
### What is mean total number of steps taken per day?
```{r, echo=TRUE}
#library (lubridate)
cleandata <- subset(data, !is.na(steps))
dtotal <- ddply(cleandata,~date,summarise,total_per_day=sum(steps),mean=mean(steps),median=median(steps))
```
Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
 
 hist(dtotal$total_per_day , main="Total number of steps taken each day",xlab="steps taken each day")
```

Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
head(dtotal) 
```

### What is the average daily activity pattern?
```{r, echo=TRUE}
dtotal2 <- ddply(cleandata,~interval ,summarise,total_per_interval=sum(steps),mean=mean(steps))
attach(dtotal2) 
plot(interval,mean ,type="o",col="red")
maxinterval = dtotal2[mean==max(dtotal2$mean), ][1]
print(maxinterval)
```

### Imputing missing values
 total number of missing values in the dataset 
```{r, echo=TRUE}
nadata <- subset(data, is.na(steps))
maxrec <- cleandata[cleandata$steps==max(cleandata$steps),]
nacount<- lapply(nadata , length)
nacount<- nacount$steps
 print(nacount)
```
 filling in all of the missing values in the dataset
 Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
stepMean <- mean(cleandata$steps)
data2<- data
data2[is.na(data2$steps),1] <- stepMean 
head(data2)
```


### Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}

  wd1 <- wday(as.Date(data2$date,format("%Y-%m-%d")) )

data2<- mutate(data2,   weekend = ifelse( (wd1 == 1 | wd1== 7) , "weekend","weekday" ) )
#png("plotN.png", width=480, height=480 )
attach(data2)
 
 qplot(data2,x=interval,y=steps ,facets=.~ weekend)

```

 